---
title: "Taxonomy_FurtherAnalysis"
author: "Alecia Gee"
date: "2025-09-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
```

```{r import data}
pfam_data <- read_csv("~/PhD/Experimental/Project 2/R_DataVisualisation/fulltable.csv")
pfam_data
```


```{r cleanup}
cleaned_data <- pfam_data %>%
  select(-contains("sequence"),
         -contains("accession"),
         -contains("Unnamed: 0"),
         -contains("lineage_and_ranks"),
         -contains("taxon")
  ) %>%
  filter(acid_alc != "-T" & acid_alc != "--") %>%
  filter(kingdom != "NA")
```

```{r unique kingdoms}
unique_kingdoms <- cleaned_data %>%
  group_by(kingdom) %>%
  summarise(group_count = n(), .groups = "drop_last")
print(unique_kingdoms)
```
Going to combine the kingdoms where n < 300. Therefore, Bamfordvirae, Heunggongvirae, Nanobdellati, Thermoproteati and Thermotogati and "Methanobacteriati".

```{r}
grouped_data_allkingdoms <- cleaned_data %>%
  group_by(acid_alc, kingdom) %>%  
  summarise(group_count = n(), .groups = "drop_last") %>%
  mutate(
    total_count = sum(group_count),
    percentage = (group_count / total_count) * 100
  ) %>%
  arrange(acid_alc, kingdom) %>%
  separate(acid_alc, into = c("acid", "alc"), sep = 1, remove = FALSE)

grouped_data_allkingdoms
```


```{r Combining small groups to other}
kingdoms_condensed <- cleaned_data %>%
  mutate(king_cond = ifelse(kingdom %in% c("Bamfordvirae", "Heunggongvirae", "Nanobdellati", "Thermoproteati", "Thermotogati", "Methanobacteriati"), "Other", kingdom))
```


```{r acid and alcohol combo}
grouped_data <- kingdoms_condensed %>%
  group_by(acid_alc, king_cond) %>%  
  summarise(group_count = n(), .groups = "drop_last") %>%
  mutate(
    total_count = sum(group_count),
    percentage = (group_count / total_count) * 100
  ) %>%
  arrange(acid_alc, king_cond) %>%
  separate(acid_alc, into = c("acid", "alc"), sep = 1, remove = FALSE)

grouped
```

```{r adding groups}
analysed_data <- grouped_data %>%
    mutate(position1_groups = case_when(
    acid %in% c("E", "D") ~ "acid",
    acid %in% c("R", "H", "K") ~ "charged",
    acid %in% c("S", "T") ~ "alcohol",
    acid %in% c("N", "Q") ~ "amide",
    acid %in% c("V", "A", "G", "I", "L", "F") ~ "aliphatic",
    acid %in% c("C", "Y", "W", "P", "M") ~ "special",
  )) %>%
    mutate(position2_groups = case_when(
    alc %in% c("E", "D") ~ "acid",
    alc %in% c("R", "H", "K") ~ "charged",
    alc %in% c("S", "T") ~ "alcohol",
    alc %in% c("N", "Q") ~ "amide",
    alc %in% c("V", "A", "G", "I", "L", "F") ~ "aliphatic",
    alc %in% c("C", "Y", "W", "P", "M") ~ "special",
  ))
```

```{r amino acid grouping to output}

groups <- c("acid", "alcohol", "charged", "amide", "aliphatic", "special")

# make all 36 combos
combos <- expand.grid(position1_groups = groups,
                      position2_groups = groups,
                      stringsAsFactors = FALSE)

# generate filtered data for each combo
results <- pmap(combos, function(position1_groups, position2_groups) {
  analysed_data %>%
    filter(position1_groups == !!position1_groups,
           position2_groups == !!position2_groups)
})

# give each element a name like "acid_alcohol"
names(results) <- paste(combos$position1_groups, combos$position2_groups, sep = "_")

# write each element of results to CSV
walk2(results, names(results), ~ write_csv(.x, paste0(.y, ".csv")))
```

Combos that are empty: alcohol_acid, charged_acid, amide_acid, special_acid, alcohol_charged, amide_charged, special_charged
There are 29 groups left.

```{r analysing acid alcohol groups}
six_groups <- read_csv("~/PhD/Experimental/Project 2/R_DataVisualisation/acidalc.csv")

heatmap(six_groups)
```

